---
title: "Marvel vs DC"
author: "Rafał Kobiela"
output: 
 html_document:
  toc: yes
  theme: cosmo
  highlight: tango
  code_folding: hide
---

# Introduction


```{r read, echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
library(readr)
library(knitr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(formattable)

heroes <- read_csv('../data/super_hero_powers_information/super_hero_powers_information.csv')
powers <- read_csv('../data/super_hero_powers_information/super_hero_powers.csv')
```
## Inspecting the Datasets {.tabset .tabset-fade .tabset-pills}

###Heroes
```{r head_heroes}
knitr::kable(head(heroes,10),caption="Heroes List (first 10 rows)")
```

###Heroes Summary
```{r summary_heroes}
summary(heroes)
```

###Heroes Structure
```{r glimpse_heroes}
glimpse(heroes)
```

###Powers
```{r head_powers}
knitr::kable(head(powers,10),caption="Superpowers List of each Heroes (first 10 rows)")
```

###Powers Summary
```{r summary_powers}
summary(powers)
```

###Powers Structure
```{r glimpse_powers}
glimpse(powers)
```

## Preparing the Data

```{r prepare,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
heroes$Weight[heroes$Weight == "-99"] <- NA
heroes$Height[heroes$Height == "-99"] <- NA

powers[,-1] <- lapply(powers[,-1], as.logical)
heroes$Gender <- as.factor(heroes$Gender)
heroes$Alignment <- as.factor(heroes$Alignment)

heroes$Gender[heroes$Gender == "-"] <- NA
heroes$`Eye color`[heroes$`Eye color` == "-"] <- NA
heroes$Race[heroes$Race == "-"] <- NA
heroes$`Hair color`[heroes$`Hair color` == "-"] <- NA
heroes$`Skin color`[heroes$`Skin color` == "-"] <- NA
heroes$Alignment[heroes$Alignment == "-"] <- NA

heroes <- heroes[,-1]
```

## Inspecting the Tidy Datasets {.tabset .tabset-fade .tabset-pills}

###Heroes
```{r head_heroes_tidy}
knitr::kable(head(heroes,10),caption="Heroes List (first 10 rows)")
```

###Heroes Summary
```{r summary_heroes_tidy}
summary(heroes)
```

###Heroes Structure
```{r glimpse_heroes_tidy}
glimpse(heroes)
```

###Powers
```{r head_powers_tidy}
knitr::kable(head(powers,10),caption="Superpowers List of each Heroes (first 10 rows)")
```

###Powers Summary
```{r summary_powers_tidy}
summary(powers)
```

###Powers Structure
```{r glimpse_powers_tidy}
glimpse(powers)
```

---


# Marvel vs DC Comics: Universe Comparison

Since most of the data consist of superheroes from either **DC Comics** or **Marvel**, I'll compare only two of them.

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
marvel_dc <- heroes %>%
    filter(Publisher %in% c("Marvel Comics","DC Comics")) 
```

## Which Superheroes have more variety of superpowers?

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
powers <- powers %>%
    mutate(total_powers = rowSums(.[2:ncol(powers)]))
    
top20_heroes <- marvel_dc %>%
    inner_join(powers, by = c('name' = 'hero_names')) %>%
    select(name, total_powers, Publisher) %>%
    arrange(desc(total_powers)) %>%
    head(20)

top20_heroes %>%
    ggplot(aes(x=reorder(name,total_powers), y=total_powers, fill=Publisher)) +
    geom_bar(stat = "identity") + 
    geom_text(aes(label = total_powers), hjust = -0.4, size = 3, color = "black") +
    labs(title = "Superheroes having most number of powers", x='Name',y='Number of Superpowers') +
    theme_bw() +
    theme(plot.title = element_text(size=17,hjust = -0.4, face = "bold"), legend.position = "bottom", legend.title=element_blank()) +
    scale_fill_manual(values = c("#377EB8", "#E41A1C")) +
    coord_flip()
```

## Which Superheroes have unique powers?

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
gathered_powers2 <- gather(powers, power, power_count, -hero_names, -total_powers)

gathered_powers2 %>%
    group_by(power) %>%
    mutate(how_many = sum(power_count)) %>%
    ungroup() %>%
    filter(how_many == 1 & power_count == 1) %>%
    inner_join(marvel_dc, by = c('hero_names' = 'name')) %>%
    select(hero_names, power, total_powers, Publisher) %>%
    arrange(desc(total_powers)) %>%
    formattable()
```

## Common Powers of Superheroes

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
heroes$Publisher[is.na(heroes$Publisher)] <- 'Other Comics'
heroes$Publisher[! (heroes$Publisher %in% c('Marvel Comics','DC Comics'))] <- 'Other Comics'

gathered_powers3 <- heroes %>%
    select(name, Publisher) %>%
    inner_join(powers, by = c('name' = 'hero_names')) %>%
    gather(power, power_count, -name, -Publisher)

top_superpowers <- gathered_powers3 %>%
    group_by(power, Publisher) %>%
    summarise(powers_used_by = sum(power_count)) %>%
    arrange(desc(powers_used_by)) %>%
    filter(powers_used_by < 800) %>%
    group_by(power) %>%
    mutate(powers_in_total = sum(powers_used_by)) %>%
    ungroup()

top_superpowers$Publisher <- factor(top_superpowers$Publisher, levels = c('Other Comics','DC Comics','Marvel Comics')) 

top_superpowers %>%
    filter(powers_in_total >= 75) %>%
    ggplot(aes(x = reorder(power,powers_in_total) , y = powers_used_by, fill = Publisher)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("#33A02C", "#377EB8", "#E41A1C")) +
    labs(title = "Common Powers of Superheroes", x = "Powers", y = "Number of Superheroes") +
    theme_bw() +
    theme(legend.position = "bottom", legend.title=element_blank(), plot.title = element_text(size=17,hjust = -0.4, face = "bold")) +
    ylim(0, max(top_superpowers$powers_used_by + 200)) +
    coord_flip()
```

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
top_superpowers %>%
    filter(powers_in_total >= 100) %>%
    ggplot(aes(x = reorder(power,powers_in_total) , y = powers_used_by, fill = Publisher)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("#33A02C", "#377EB8", "#E41A1C")) +
    labs(title = "Common Powers of Superheroes", x = "Powers", y = "Number of Superheroes") +
    theme_bw() +
    theme(legend.position = "bottom",legend.title=element_blank(), plot.title = element_text(size=17,hjust = -0.4, face = "bold")) +
    coord_flip()
```

## Gender Distribution in each universe

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
ggplot(data=subset(heroes, !is.na(Gender)), aes(x = Publisher, fill = Gender)) +
    geom_bar(position = "fill") +
    scale_fill_brewer(palette = "Paired") +
    labs(x = "", y = "") +
    theme_bw()
```

## Good & Bad Alignment in each universe

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
heroes$Alignment <- factor(heroes$Alignment, levels = c('bad','neutral','good'))

ggplot(data=subset(heroes, !is.na(Alignment)), aes(x = Publisher, fill = Alignment)) +
    geom_bar(position = "fill") +
    scale_fill_brewer(palette = "Set2") +
    labs(x = "", y = "") +
    theme_bw()
```

## Common Races of Superheroes

```{r ,echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
races <- heroes %>%
    filter(!is.na(Race)) %>%
    group_by(Publisher, Race) %>%
    summarise(race_total = n()) %>%
    arrange(desc(race_total))

races <- races %>%
    group_by(Race) %>%
    mutate(total = sum(race_total)) %>%
    ungroup()

races$Publisher <- factor(races$Publisher, levels = c('Other Comics','DC Comics','Marvel Comics'))

races %>%
    filter(total >= 5) %>%
    ggplot(aes(x = reorder(Race,total) , y = race_total, fill = Publisher)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("#33A02C", "#377EB8", "#E41A1C")) +
    labs(title = "Common Races of Superheroes", x = "Races", y = "Number of Superheroes") +
    theme(legend.position = "bottom", legend.title=element_blank()) +
    scale_y_continuous(breaks = seq(0,100,5)) +
    coord_flip()
```
